---
title: "SEKI Data Analysis"
author: "Ethan Tenison"
date: "`r format(Sys.Date(), '%B %d, %Y') `"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_depth: 3
    collapsed: yes
    number_sections: yes
    fig_caption: yes
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tidyverse) #Data cleaning and visualizations
library(janitor) #Basic data preprocessing 
library(readr) #Reading data from various formats 
library(fontawesome) #Icons

library(visNetwork) #visualizing network data using html 
library(igraph) #Barebones network analysis package 

library(leaflet) #html/javascript mapping
library(sf) #reading shapefiles 
library(sp) #another package for spatial processing
library(htmlwidgets) #for saving leaflet 

```


# Introduction


# Full Network

Visualizing the preliminary dataset using `Visnetwork`is simple given the dataset has already been cleaned. 

```{r message=FALSE, warning=FALSE}



#Reading in the data
node_attri <-
  read_csv("data/processed/SEKI Node Attributes.csv") |> clean_names()
edges <-
  read_csv("data/processed/SEKIedgelist.csv") |> clean_names()

#Converting to network type
g1 <- graph_from_data_frame(d = edges, directed = TRUE)

gvis <- toVisNetworkData(g1)

nodes <- gvis$nodes
nodes <- nodes |> mutate(color = "#EE4B2B")

network <-
  visNetwork(nodes,
             gvis$edges,
             main = "SEKI PACE Network",
             width = "100%",
             height = "850px") |>
  visEdges(
    smooth = T,
    arrows = list(
      to = list(enabled = TRUE, scaleFactor = .5),
      width = 3
    ),
    color = list(highlight = "black")
  ) |>
  visNodes(color = list(
    background = "white",
    border = "black",
    highlight = list(background = "#A9A9A9", border = "black"),
    hover = list(background = "#A9A9A9", border = "black")
  )) |>
  visPhysics(stabilization = FALSE)  |>
  visIgraphLayout(
    smooth = FALSE,
    physics = FALSE,
    layout = "layout_with_fr",
    randomSeed = 27
  ) |>
  visInteraction(navigationButtons = TRUE) 

network 
```

# Land Management Only 

Next we're going to look at only those entities that have a defined geographic location. In addition, the size is set to equal the total number of connections. By far, the `Inyo National Forest Supervisors Office` has the most connections at 41. 

```{r landmgr_only}

landmgr <- node_attri |>
  select(name, landmgr) |>
  filter(landmgr == 1)

joined <- edges |>
  left_join(landmgr, by = c("ego" = "name")) |>
  filter(landmgr == 1)

#Converting to network type
g1 <- graph_from_data_frame(d = joined, directed = TRUE)
TotalDegree <- igraph::degree(g1)
gvis <- toVisNetworkData(g1)

#nodes
nodes <- gvis$nodes
nodes <- nodes |>
  left_join(node_attri, by = c("id" = "name"))
nodes$value <- TotalDegree[match(nodes$id, names(TotalDegree))]
nodes$label <- paste(nodes$id, "\n", nodes$value)
nodes$color[nodes$orgtype == "State Agency"] <- "blue"
nodes$color[nodes$orgtype == "Federal Agency"] <- "red"
nodes$color[nodes$orgtype == "Tribal"] <- "green"
nodes$color[nodes$orgtype == "NGO"] <- "yellow"
nodes$color[nodes$orgtype == "Local agency"] <- "orange"
nodes$color[nodes$orgtype == "Private"] <- "purple"
nodes$color[is.na(nodes$orgtype)] <- "gray"


network <-
  visNetwork(nodes,
             gvis$edges,
             main = "SEKI PACE Network (Landbased only)",
             width = "100%",
             height = "850px") |>
  visEdges(
    smooth = T,
    arrows = list(
      to = list(enabled = TRUE, scaleFactor = .5),
      width = 3
    ),
    color = list(highlight = "black")
  ) |>
  visNodes(color = list(
    background = "white",
    border = "black",
    highlight = list(background = "#A9A9A9", border = "black"),
    hover = list(background = "#A9A9A9", border = "black")
  )) |>
  visPhysics(stabilization = FALSE)  |>
  visIgraphLayout(
    smooth = FALSE,
    physics = FALSE,
    layout = "layout_with_fr",
    randomSeed = 27
  ) |>
  visInteraction(navigationButtons = TRUE) |>
  addFontAwesome(name = "font-awesome-visNetwork") |>
  visLegend(
    position = "left",
    addNodes = list(
      list(
        label = "State Agency",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "blue")
      ),
      list(
        label = "Federal Agency",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "red")
      ),
      list(
        label = "Tribal",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "green")
      ),
      list(
        label = "NGO",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "yellow")
      ),
      list(
        label = "Local Agency",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "orange")
      ),
      list(
        label = "Private",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "purple")
      ),
      list(
        label = "Unknown",
        shape = "icon",
        icon = list(code = "f111", size = 20, color = "gray")
      )
      
    ),
    useGroups = FALSE,
    stepY = 50
  )




network 

```

# Ploting the boundaries 

The raw shapefile includes nearly 1,500 different polygons, which make it unusable in the current state. 

```{r firstmap}

bg_raw <- st_read("data/external/SEKI_PACE.shp", stringsAsFactors = FALSE)
bg_raw <-st_transform(bg_raw,"+proj=longlat +ellps=WGS84 +datum=WGS84")

#I deselected the geometry column to I could easily analyze the data. It's computationally expensive otherwise.
bg <-
  bg_raw |> dplyr::select(
    "Category",
    "Own_Type",
    "Own_Name"   ,
    "Loc_Own"    ,
    "Mang_Type"  ,
    "Mang_Name" ,
    "Loc_Mang" ,
    "Des_Tp" ,
    "Loc_Ds", 
    "Unit_Nm"  ,
    "Loc_Nm"
  ) |> 
  st_set_geometry(NULL)

bg <- as.data.frame(bg)

write.csv(bg, "data/interim/converted_shapefile.csv")

#Quick visualization of the raw shapefile 
leaflet() |> 
  addTiles() |> 
  addPolygons(data = bg_raw)

```

# Joining the polygons to the SEKI point data 

The shapefile of SEKI jurisdictions is very messy. After repeated attempts to filter out unused polygons, it became clear that they are overlapping and broken up into pieces, making it difficult to look at regardless of the filtering. In addition, of those nodes that have lonlat values, the majority lie outside of the polygons.  

```{r spatial_joining}



#imputing avg long lat values to na's, to be changed later 
avg_lat <- mean(nodes$centroid_lat, na.rm = TRUE)
avg_lon <- mean(nodes$centroid_lon, na.rm = TRUE)

nodes_sf <- nodes
nodes_sf$centroid_lat[is.na(nodes_sf$centroid_lat)] <- avg_lat
nodes_sf$centroid_lon[is.na(nodes_sf$centroid_lon)] <- avg_lon

nodes_sf = st_as_sf(nodes_sf, coords = c("centroid_lon", "centroid_lat"), 
                 crs = 4326, agr = "constant")

#Quick viz of points
leaflet() |> 
  addTiles() |> 
  addCircles(data = nodes_sf)

#My attempts at filtering out polygons 
bg <- bg_raw |> select("Loc_Mang", "Unit_Nm","GIS_Acres", "geometry")
out <- st_intersection(bg, nodes_sf)
gis_acres <- select(out, GIS_Acres)
bg <- as.data.frame(bg)
polygons <- gis_acres |> left_join(bg, by = "GIS_Acres") |> distinct()
polygons <- polygons |> select("GIS_Acres",  "Loc_Mang",   "Unit_Nm","geometry.y") |> rename("geometry" = "geometry.y")
polygons$geometry.x <- NULL

polygons <- st_sf(polygons)
#Quick viz of points
leaflet() |> 
  addTiles() |> 
  addPolygons(data = bg_raw, label = bg_raw$Unit_Nm) |> 
  addCircles(data = nodes_sf, fillColor = "red", color = "red", label = nodes_sf$id)


```


# Leaflet Map with ring of undefined points  


```{r leaflet_with_ring}


#Quick viz of points
leaflet() |>
  addTiles(urlTemplate = 'https://tiles.stadiamaps.com/tiles/outdoors/{z}/{x}/{y}{r}.png',
           attribution = '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors') |>
  addCircles(
    data = nodes_sf,
    color = "red",
    weight = nodes_sf$value,
    label = nodes_sf$id
  ) 


#assigning ego and alter coordinates based on node info 
coords <- nodes |>
  select(id,centroid_lat,centroid_lon)

edges <- joined |>
  left_join(coords, by = c("ego" = "id")) |>
  left_join(coords, by = c("alter" = "id")) |> 
  dplyr::mutate(centroid_lat.y = replace_na(centroid_lat.y, avg_lat),
                centroid_lon.y = replace_na(centroid_lon.y, avg_lon))

edges <- data.frame(group = c(edges$ego, edges$alter),
                    lat = c(edges$centroid_lat.x, edges$centroid_lat.y),
                    lon = c(edges$centroid_lon.x, edges$centroid_lon.y)
                    )


#Quick viz of points
leaflet() |>
  addTiles(urlTemplate = 'https://tiles.stadiamaps.com/tiles/outdoors/{z}/{x}/{y}{r}.png',
           attribution = '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors') |>
  addCircles(
    data = nodes_sf,
    color = "red",
    weight = nodes_sf$value,
    label = nodes_sf$id
  ) |> 
  addPolylines(data = edges, lng = ~lon, lat = ~lat, group = ~group, weight = 1)


```


# Reassigning coords to nodes without lonlat values

In order to show all of the organizations, even those without a physical address, I created a ring 

```{r the_ring}

for_ring <- nodes[is.na(nodes$centroid_lat),]

number_of_orgs <- nrow(for_ring)

outer_circle <- for_ring |> 
  dplyr::mutate(centroid_lat = replace_na(centroid_lat, avg_lat),
                centroid_lon = replace_na(centroid_lon, avg_lon)) |> 
  dplyr::mutate(
    x = cos((row_number() - 1) / number_of_orgs * 2 * pi),
    y = sin((row_number() - 1) / number_of_orgs * 2 * pi)
  ) |> 
  dplyr::mutate(
    new_lat = centroid_lat + y*1.5,
    new_lon = centroid_lon + x*1.5
  )

outer_circle_toadd <- outer_circle |>
  select(id, new_lon, new_lat) |> 
  mutate(location = "Nonphysical")

outer_circle = st_as_sf(outer_circle, coords = c("new_lon", "new_lat"), 
                 crs = 4326, agr = "constant")

leaflet() |>
  addTiles(urlTemplate = 'https://tiles.stadiamaps.com/tiles/outdoors/{z}/{x}/{y}{r}.png',
           attribution = '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors') |>
  addCircles(
    data = outer_circle,
    color = "red",
    weight = outer_circle$value,
    label = outer_circle$id
  )


  
```

# Adding the ring

```{r ring_add}
#Joining the ring with the physical location nodes

original_value <- nodes_sf |> select(id, value)

nodes_sf <- nodes |> 
  left_join(outer_circle_toadd) |> 
  mutate(
    centroid_lon = ifelse(is.na(centroid_lon), new_lon, centroid_lon),
    centroid_lat = ifelse(is.na(centroid_lat), new_lat, centroid_lat),
    location = ifelse(is.na(location), "Physical", location),
    value = value +10
  )

write_rds(nodes_sf, "data/processed/nodes_4_shiny.rds")

coords <- nodes_sf |>
  select(id,centroid_lat,centroid_lon)
nodes_sf= st_as_sf(nodes_sf, coords = c("centroid_lon", "centroid_lat"), 
                 crs = 4326, agr = "constant")

write_rds(coords, "data/processed/coords_4_shiny.rds")

#assigning ego and alter coordinates based on node info 
edges <- joined |>
  left_join(coords, by = c("ego" = "id")) |>
  left_join(coords, by = c("alter" = "id")) 

write_rds(edges, "data/processed/edges_4_shiny.rds")

edges <- data.frame(group = c(edges$ego, edges$alter),
                    lat = c(edges$centroid_lat.x, edges$centroid_lat.y),
                    lon = c(edges$centroid_lon.x, edges$centroid_lon.y)
                    )



#Assigning a palette 
nodes_sf$location <- as.factor(nodes_sf$location)
pal <- colorFactor(
  palette = c('red', 'blue'),
  domain = nodes_sf$location
)


#Leaflet map with ring added 


m <- leaflet(nodes_sf) |>
  # addTiles(urlTemplate = 'https://tiles.stadiamaps.com/tiles/outdoors/{z}/{x}/{y}{r}.png',
  #          attribution = '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors') |>
  addProviderTiles(providers$OpenStreetMap.DE) |> 
  addCircles(
    color = ~pal(nodes_sf$location),
    weight = nodes_sf$value,
    label = nodes_sf$id
  ) |> 
  addPolylines(data = edges, lng = ~lon, lat = ~lat, group = ~group, weight = 1) |> 
  addLegend("bottomright",
            pal = pal,
            values = ~nodes_sf$location,
            title = "")


saveWidget(m, file="reports/ring_map.html")
```